---
title: "Practice - R for Data Science"
author: "Ben Nickley"
date: "December 7, 2017"
output: pdf_document
---
```{r}
library(ggplot2)
```
```{r}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# Ch 3 - Data Visualization

[R for Data Science - Ch 3](http://r4ds.had.co.nz/data-visualisation.html)

ggplot creates the coordinate system w/ the first arguemnt, which tells ggplot what data you are using

* add layers with subsequent arguemnts (i.e. geom_point)
```{r echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = 'red', size = 2)
```

This is the ggplot skeleton:
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
 
```{r} 
ggplot(data = mpg) +
geom_point(mapping = aes(x = class, y = drv))

ggplot(data = mpg) +
geom_point(mapping = aes(x = hwy, y = cyl))
```

Map 'class' of car to different aesthetics (aes)
```{r Plots - Experimenting with aes parameters}
# Left
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy, fill = class), shape = 24, color = 'black', alpha = 0.75)

# Middle
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Right
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r Mapping continuous variables in aes}
ggplot(data = mpg) +
  geom_point(mapping = aes(x=cyl, y=displ, color=hwy), stroke=2)

# What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?
ggplot(data = mpg) +
  geom_point(mapping = aes(x=cyl, y=hwy, colour = displ < 5), stroke=2)
```









